---
title: "Using fonts in R graphics with knitr"
author: "Yixuan Qiu"
date: "07/21/2014"
output:
  html_document:
    keep_md: yes
---

Thanks to the [issue report](https://github.com/yihui/knitr/issues/799) by
[yufree](https://github.com/yufree) and Yihui's
[kind work](https://github.com/yihui/knitr),
from version 1.6.10 (development version), **knitr** starts to support using
[**showtext**](https://github.com/yixuan/showtext)
to change fonts in R plots. To demonstrate its usage, this document
itself serves as an example.

We first do some setup work, mainly about setting options that control
the appearance of the plots. Notice that if you create plots in PNG
format (the default format for HTML output), it is strongly recommended
to use the `CairoPNG` device rather than the default `png`, since
the latter one could produce quite ugly plots when using **showtext**.

```{r setup}
knitr::opts_chunk$set(dev="CairoPNG", fig.width=7, fig.height=7, dpi = 72)
options(digits = 4)
```

Then we can load **showtext** package and add fonts to it. Details about
font loading are explained in the
[introduction document](https://github.com/yixuan/showtext/blob/master/README.md)
of **showtext** and also in the help topic `?sysfonts::font.add`.
While searching and adding fonts may be a tedious work,
package **sysfonts** (which **showtext** depends on)
provides a convenient function `font.add.google()` to automatically download
and use fonts from the Google Fonts repository
([http://www.google.com/fonts](http://www.google.com/fonts)).
The first parameter is the font name in Google Fonts and the second one is
the family name that will be used in R plots.

```{r fonts, message=FALSE}
library(showtext)
font.add.google("Lobster", "lobster")
```

After adding fonts, simply set the `fig.showtext` option in the code block
where you want to use **showtext**, and then specify the family name you
just added.

```{r fig.showtext=TRUE, fig.align='center'}
plot(1, pch = 16, cex = 3)
text(1, 1.1, "A fancy dot", family = "lobster", col = "steelblue", cex = 3)
```

